---
title: "Timing of breeding reveals a trade-off between constitutive immune investment and life history in a migratory bird"
author: "David Chang van Oordt, Conor C. Taff, Thomas A. Ryan, Maren N. Vitousek"
bibliography: vitousek2018.bib
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE,
                      fig.align = "center")

# Load packages.
library(lme4)
library(bbmle)
library(gridExtra)
library(dagitty)
library(lubridate)
library(viridis)
library(here)
library(sjPlot)
library(gtsummary)
library(MASS)
library(tidyverse)
library(knitr)
library(broom)
library(lmerTest)
library(here)

# over dispersion
dfun <- function(object) {
  with(object,sum((weights * residuals^2)[weights > 0])/df.residual)
  }

# Read data files:
# Assay data
bk <- read_csv(here("data", "bk.csv"), col_types = 
                 cols(
                   Sex = col_factor(),
                   bleed.type = col_factor()
                 ))

# Physio and fitness data + bka
df <- read_csv(here("data", "df.csv"), 
               col_types = cols(
                 Sex = col_factor(),
                 Experiment = col_factor(),
                 Treatment = col_factor(),
                 Age = col_factor(),
                 Plasma_Analyzed = col_factor(),
                 Nest_Fate = col_factor(),
                 Treatment2 = col_factor()
               ))

# Nestling data
nd <- read_csv(here("data", "nd.csv"), 
               col_types = cols(
                 Sex = col_factor(),
                 Experiment = col_factor(),
                 Treatment = col_factor(),
                 Age = col_factor(),
                 Plasma_Analyzed = col_factor(),
                 Nest_Fate = col_factor() 
                 ))

# Daily feeding visits
rfid_d <- read_csv(here("data", "feed_by_day.csv") )

# Hourly feeding data
pc <- read_csv(here("data", "pc.csv"),
               col_types = cols(
                 Experiment = col_factor(),
                 Treatment = col_factor(),
                 f_age = col_factor(),
                 Plasma_Analyzed = col_factor(),
                 Nest_Fate = col_factor()
                 ))
```

# Method assessment

## Assay repeatability

```{r repeatability, fig.width = 8, fig.height = 4}
# Repeatability by plate

pseurep <- bk %>%
  ggplot() + geom_point(aes(x=rep1, y=rep2)) +
  geom_abline(slope =1) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none") +
  labs(title="Pseureplicates - Plate effect",
       x = "No. of CFUs in Plate 1",
       y = "No. of CFUs in Plate 2") +
  scale_y_continuous(expand = c(0,0), limits = c(0,700)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,700)) +
  annotate("text", x = 500, y = 600, 
           label = paste("r","=",
                         round(cor(cbind(bk$rep1, bk$rep2),
                                   method = "pearson")[1,2], 3) ) )

# Repeatability by reaction

reps <- inner_join(filter(bk, rxn.rep == 1) %>% select(Individual_Band, bkc),
           filter(bk, rxn.rep == 2) %>% select(Individual_Band, bkc),
           by = "Individual_Band")

rxnrep <- reps %>%
  ggplot() + geom_point(aes(x=bkc.x, y=bkc.y)) +
  geom_abline(slope =1) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none") +
  labs(title="Reaction repeatibility",
       x = "Prop. Bacteria Killed in Challenge 1",
       y = "Prop. Bacteria Killed in Challenge 2") +
  annotate("text", x = 0.5, y = 0.95, 
           label = paste("r =",round(cor(reps[,-1])[1,2], 3) ) ) 

grid.arrange(pseurep, rxnrep, ncol = 2)
```

## Treament type comparisons

```{r treatment comparison}
# Boxplot
df %>% 
  mutate(timing = as.factor(
    ifelse(Experiment == "Water", "30 minutes", "<3 minutes")
  )) %>% 
  ggplot() +
  geom_boxplot(aes(x = Treatment2, y = Bacteria_Killing_Assay),
               size = 1, width = 0.25, fill = "grey") +
  geom_point(aes(x = Treatment2, y = Bacteria_Killing_Assay), 
             position = position_jitter(width = 0.1)) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none",
        legend.title.align = 0.5,
        legend.background = element_rect(fill=NA),
        axis.text.x=element_text(angle=20,hjust=1)) +
  scale_x_discrete(labels = c("Predator-Dull", "Predator-Control", "Long-term Control", "Control-Control", "Control-Dull")) +
  labs(x = "Timing of Blood Sample",
       y = "Prop. of Bacteria Killed")

# T-test
bka_mod1 <- lm(bk ~ Treatment2, data = df)
summary(aov(bk ~ Treatment2, data = df))
```
# Results

## BKA in our population

```{r bka summary}
bk %>%
  group_by(Individual_Band) %>% 
  summarise(bkc = mean(bkc)) %>%
  ggplot() + geom_histogram(aes(bkc), bins = 15, colour = "white") +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(limits = c(0, 10), breaks=seq(0,10,by=2),
                     expand = c(0,0)) +
  geom_vline(xintercept=0, linetype = "dashed", colour = "grey") +
  labs(x = "Prop. Bacteria Killed", y = "No. of individuals")

summ_bk <- group_by(bk, Individual_Band) %>% summarise(bkc = mean(bkc))
summ <- matrix(0, nrow = 1, ncol = 6)
colnames(summ) <- c("Mean", "S.D.", "2.5%", "97.5%", "Min.", "Max.")
summ[1,1] <- mean(summ_bk$bkc, na.rm = TRUE)
summ[1,2] <- sd(summ_bk$bkc, na.rm = TRUE)
summ[1,3] <- summ[1,1] - 1.96*summ[1,2]/sqrt(nrow(summ_bk))
summ[1,4] <- summ[1,1] + 1.96*summ[1,2]/sqrt(nrow(summ_bk))
summ[1,5] <- min(summ_bk)
summ[1,6] <- max(summ_bk)
kable(summ)
```

## BKA and Reproductive Effort

### General trend evaluation

#### BKA and age

```{r BKA age}
summary(glm(bk ~ Age,  family=Gamma(link="inverse"),
            data = df))
```

#### BKA and Lay Date

```{r bka laydate}
summary(glm(bk ~ ylaydate, family = Gamma(link="inverse"),
                  data = df))
```

### Clutch Size

#### Clutch size data distribution

```{r clutch distribution}
ggplot(df) +
  geom_histogram(aes(x=Clutch_Size), colour = "white", bins = 15) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x= "Clutch size",
       y = "No. of individuals") +
  scale_x_continuous(limits = c(0,8),
                                    breaks = 1:8)
```

#### Clutch size models

##### Null model

```{r clutch null}
clutch_m1 <- glm(Clutch_Size ~ Age + ylaydate,
                family = quasipoisson(link = "log"), 
                data = df)
summary(clutch_m1)
```

##### BKA model

```{r clutch BKA}
clutch_m2 <- glm(Clutch_Size ~ Bacteria_Killing_Assay + Age + ylaydate,
                family = quasipoisson(link = "log"), 
                data = df)
summary(clutch_m2)
```

##### Interaction model

```{r clutch interaction}
clutch_m3 <- glm(Clutch_Size ~ Bacteria_Killing_Assay*ylaydate + Age,
                 family = quasipoisson(link = "log"), 
                 data = df)
summary(clutch_m3)
```

##### Model comparison

```{r aic clutch}
# redefine models to poisson models
clutch_m1p <- glm(Clutch_Size ~ Age + ylaydate,
                family = poisson(link = "log"), 
                data = df)
clutch_m2p <- glm(Clutch_Size ~ Bacteria_Killing_Assay + Age + ylaydate,
                family = poisson(link = "log"), 
                data = df)
clutch_m3p <- glm(Clutch_Size ~ Bacteria_Killing_Assay*ylaydate + Age,
                 family = poisson(link = "log"), 
                 data = df)

ICtab(clutch_m1p, clutch_m2p, clutch_m3p, dispersion = dfun(clutch_m1p), type = "qAICc")
```

### Nestling Feeding Rate

#### Feeding Rate variation

```{r visits variation}

# Note that Offset refers to Nestling Age.

ggplot(rfid_d) + 
  geom_point(aes(x= Offset, y = FemFeed, group = f_rfid), alpha = 0.1, show.legend = F) + 
  geom_line(aes(x= Offset, y = FemFeed, group = f_rfid), alpha = 0.1, show.legend = F) +
  geom_boxplot(aes(x = Offset, y = FemFeed, group = as.factor(Offset), fill = as.factor(Offset)),
               width = 0.3) +
  scale_x_continuous(expand = c(0,0), limits = c(0,20)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none") +
  labs(title="Daily provisioning visits per female",
       x = "Nestling Age (days)",
       y = "Number of visits per day")
```

#### Feeding Rate models

```{r data scaling}
scaled_pc <- pc %>%
  mutate(scBrood_Size_Hatching = scale(Brood_Size_Hatching),
         scOffset = scale(Offset),
         scHour = scale(Hour),
         scjlaydate = scale(jlaydate),
         sctemp = scale(temp),
         scMalFeed = scale(MalFeed),
         scbkc = scale(bkc),
         Individual_Band = as.factor(Individual_Band))
```

##### For all nests

###### Null model

These models are modified from [@vitousekLingeringImpactStress2018]

```{r feeding null}
provisioning_m1 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band),
                         family=poisson(link="log")); 
summary(provisioning_m1)

```

###### BKA model

```{r feeding BKA}
# bkc model
provisioning_m2 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band) + scbkc,
                         family=poisson(link="log"))
summary(provisioning_m2)
```

###### Interaction model

```{r feeding interaction}
provisioning_m3 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                            I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                            scHour + I(scHour^2) + sctemp + 
                            Treatment + scBrood_Size_Hatching*scOffset + 
                            (1|Individual_Band) + scbkc*scjlaydate,
                          family=poisson(link="log"))
summary(provisioning_m3)
```

###### Model comparison

```{r aic provisioning}
ICtab(provisioning_m1, provisioning_m2, provisioning_m3, type = "AICc")
```

#### Including Mate provisioning

##### Null model

```{r mate provisioning null}
provisioning_m4 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scMalFeed + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band),
                         family=poisson(link="log"),
                         control = glmerControl(optimizer = "Nelder_Mead")); 
summary(provisioning_m4)
```

##### BKA model

```{r mate provisioning bka}
provisioning_m5 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scMalFeed + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band) + scbkc,
                         family=poisson(link="log"))
summary(provisioning_m5)
```

##### Interaction model

```{r mate provisioning interaction}
provisioning_m6 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + sctemp + 
                           Treatment + scBrood_Size_Hatching*scOffset + scMalFeed +
                           (1|Individual_Band) + scbkc*scjlaydate,
                         family=poisson(link="log"))
summary(provisioning_m6)
```

###### Model Comparison

```{r mate provisioning AIC}
ICtab(provisioning_m4, provisioning_m5, provisioning_m6, type = "AIC")
```

### Mass Loss

#### Mass loss distribution

```{r mass loss histogram}
ggplot(df) +
  geom_histogram(aes(x=prop.ml),
                 colour = "white", bins = 15) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x= "Mass lost (g)",
       y = "No. of individuals")
```

#### Mass loss models

##### Null model

```{r mass loss null}
prop_mass_loss_m1 <- glm(prop.ml ~ ylaydate + Brood_Size_Hatching + Treatment2, 
                         family = gaussian(link = "identity"),
                         data = df)
summary(prop_mass_loss_m1)
```

##### BKA model

```{r mass loss bka}
prop_mass_loss_m2 <- glm(prop.ml ~ Bacteria_Killing_Assay + ylaydate + Brood_Size_Hatching + Treatment2, 
                         family = gaussian(link = "identity"),
                         data = df)
summary(prop_mass_loss_m2)
```

##### Interaction model

```{r mass loss interaction}
prop_mass_loss_m3 <- glm(prop.ml ~ Bacteria_Killing_Assay*ylaydate + Brood_Size_Hatching + Treatment2, 
                         family = gaussian(link = "identity"),
                         data = df)
summary(prop_mass_loss_m3)
```

##### Model comparison

```{r aic mass loss}
ICtab(prop_mass_loss_m1, prop_mass_loss_m2, prop_mass_loss_m3, type = "AICc")
```

## BKA and reproductive success

### Fledging success

#### Fleding success models

##### Null model
```{r success null}
success_m1 <- glm(success ~ Brood_Size_Hatching + ylaydate + Treatment2, 
                  family = binomial(link = "logit"), 
                  data = df )
summary(success_m1)
```

##### BKA model
```{r success BKA}
success_m2 <- glm(success ~ Bacteria_Killing_Assay + Brood_Size_Hatching + 
                    ylaydate + Treatment2, family = binomial(link = "logit"), 
                  data = df )
summary(success_m2)
```

##### Interaction model
```{r success interaction}
success_m3 <- glm(Success ~ Bacteria_Killing_Assay*ylaydate + 
                    Brood_Size_Hatching + Treatment2, 
                  family = binomial(link = "logit"), 
                  data = df %>%
                    mutate(Success = ifelse(Nest_Fate == "Fledged", 1, 0)))
summary(success_m3)
```

##### Model comparison
```{r success AIC}
ICtab(success_m1, success_m2, success_m3, type = "AICc")
```

### Number of nestlings fledged

#### Number fledged distribution

```{r fledged histogram}
ggplot(filter(df, .fledged > 0 )) +
  geom_histogram(aes(x=.fledged),
                 colour = "white", bins = 15) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x= "Number of nestlings fledged",
       y = "No. of individuals")
```

#### Number fledged models

##### Null model

```{r nfledged null}
fledged_m1 <- glm(.fledged ~ Brood_Size_Hatching + ylaydate +  Treatment2, 
                  family = quasipoisson(link = "log"), 
                  data = filter(df, .fledged > 0 ) )
summary(fledged_m1)
```

##### BKA model

```{r nfledged BKA}
fledged_m2 <- glm(.fledged ~ Bacteria_Killing_Assay + Brood_Size_Hatching + 
                    ylaydate + Treatment2, family = quasipoisson(link = "log"), 
                  data = filter(df, .fledged > 0 ) )
summary(fledged_m2)
```

##### Interaction model

```{r nfledged interaction}
fledged_m3 <- glm(.fledged ~ Bacteria_Killing_Assay*ylaydate + 
                    Brood_Size_Hatching + Treatment2, 
                  family = quasipoisson(link = "log"), 
                  data = filter(df, .fledged > 0 ) )
summary(fledged_m3)
```

##### Model comparison

```{r aic nfledged}
# redefine models to poisson
fledged_m1p <- glm(.fledged ~ Brood_Size_Hatching + ylaydate +  Treatment, 
                  family = poisson(link = "log"), 
                  data = filter(df, .fledged > 0 ) )
fledged_m2p <- glm(.fledged ~ Bacteria_Killing_Assay + Brood_Size_Hatching + 
                    ylaydate + Treatment, family = poisson(link = "log"), 
                  data = filter(df, .fledged > 0 ) )
fledged_m3p <- glm(.fledged ~ Bacteria_Killing_Assay*ylaydate + 
                    Brood_Size_Hatching + Treatment, 
                  family = poisson(link = "log"), 
                  data = filter(df, .fledged > 0 ) )

ICtab(fledged_m1p, fledged_m2p, fledged_m3p, dispersion = dfun(fledged_m1p), type = "qAICc")
```

### Nestling mass

#### Nestling mass distribution

```{r nestling mass histogram}
ggplot(filter(nd, Nestling_Fate == "Fledged")) +
  geom_histogram(aes(x = Nestling_Mass),
                 colour = "white", bins = 15) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x= "Nestling mass at day 12 (g)",
       y = "No. of individuals")
```

#### Nestling mass models

##### Null model

```{r nestling mass null}
nestling_m1 <- lmer(Nestling_Mass ~ Brood_Size_Hatching + 
                     ylaydate + Treatment2 + temp + (1|Unit_Box), 
                   data = filter(nd, 
                                 Nestling_Fate == "Fledged",
                                 Bacteria_Killing_Assay != "NA") )
summary(nestling_m1)
```

##### BKA model

```{r nestling mass bka}
nestling_m2 <- lmer(Nestling_Mass ~ Bacteria_Killing_Assay + 
                     Brood_Size_Hatching + temp +
                      ylaydate + Treatment2 + (1|Unit_Box), 
                   data = filter(nd, 
                                 Nestling_Fate == "Fledged",
                                 Bacteria_Killing_Assay != "NA") )
summary(nestling_m2)
```

##### Interaction model

```{r nestling mass interaction}
nestling_m3 <- lmer(Nestling_Mass ~ Bacteria_Killing_Assay*ylaydate + 
                     Brood_Size_Hatching + temp +
                      Treatment2 + (1|Unit_Box), 
                   data = filter(nd, 
                                 Nestling_Fate == "Fledged",
                                 Bacteria_Killing_Assay != "NA") )
summary(nestling_m3)
```

##### Model comparison

```{r aic neslting mass}
ICtab(nestling_m1, nestling_m2, nestling_m3, type = "AICc")
```

## Survival

### 2020 Return Rate

#### Return rate models

##### Null model
```{r return null}
return_m1 <- glm(return2020 ~ success + ylaydate + Age + Treatment2,
              family = binomial(link = "logit"),
              data = df)
summary(return_m1)
```

##### BKA model
```{r return BKA}
return_m2 <- glm(return2020 ~ success + ylaydate + Age + Treatment2 + Bacteria_Killing_Assay,
                 family = binomial(link = "logit"),
                 data = df)
summary(return_m2)
```

##### Interaction model
```{r return interaction}
return_m3 <- glm(return2020 ~ success + Age + Treatment2 + Bacteria_Killing_Assay*ylaydate,
                 family = binomial(link = "logit"),
                 data = df)
summary(return_m3)
```

##### Model comparison
```{r return AIC}
ICtab(return_m1, return_m2, return_m3, type = "AICc")
```


# References

