---
title: "Timing of breeding shapes the trade-offs between bactericidal ability and life history in a migratory bird"
author: "David Chang van Oordt, Conor C. Taff, Thomas A. Ryan, Maren N. Vitousek"
bibliography: vitousek2018.bib
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE,
                      fig.align = "center")

# Load packages.
library(lme4)
library(bbmle)
library(gridExtra)
library(dagitty)
library(lubridate)
library(viridis)
library(here)
library(sjPlot)
library(gtsummary)
library(MASS)
library(tidyverse)
library(knitr)
library(broom)
library(lmerTest)
library(here)

## functions
# legend extraction
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
# over dispersion
dfun <- function(object) {
  with(object,sum((weights * residuals^2)[weights > 0])/df.residual)
  }

# Read data files:
# Assay data
bk <- read_csv(here("data", "bk.csv"), col_types = 
                 cols(
                   Sex = col_factor(),
                   bleed.type = col_factor()
                 ))

# Physio and fitness data + bka
df <- read_csv(here("data", "df.csv"), 
               col_types = cols(
                 Sex = col_factor(),
                 Experiment = col_factor(),
                 Treatment = col_factor(),
                 Age = col_factor(),
                 Plasma_Analyzed = col_factor(),
                 Nest_Fate = col_factor(),
                 Treatment2 = col_factor()
               ))

# Nestling data
nd <- read_csv(here("data", "nd.csv"), 
               col_types = cols(
                 Sex = col_factor(),
                 Experiment = col_factor(),
                 Treatment = col_factor(),
                 Age = col_factor(),
                 Plasma_Analyzed = col_factor(),
                 Nest_Fate = col_factor() 
                 ))

# Daily feeding visits
rfid_d <- read_csv(here("data", "feed_by_day.csv") )

# Hourly feeding data
pc <- read_csv(here("data", "pc.csv"),
               col_types = cols(
                 Experiment = col_factor(),
                 Treatment = col_factor(),
                 f_age = col_factor(),
                 Plasma_Analyzed = col_factor(),
                 Nest_Fate = col_factor()
                 ))
```

# Goals and Hypotheses

The general paradigm states that immunity trades off with life history traits. 
The rationale is that any energy and resources invested on immunity comes off from other bodily functions like reproduction. 
Here we aim to look for evidence between innate immunity and life history trade-offs in Tree Swallows (TRES, *Tachycineta bicolor*).

We have 3 central hypotheses to evaluate:  

1. Immunity trades off with the physiological response to stressors
2. Immunity trades off with reproductive effort
3. Trade-offs between immunity and reproductive effort lead to costs to reproductive success

We use bactericidal ability of plasma as a measure of innate immunity. bactericidal ability using blood plasma is a complement-dependent process.

# Methods

We performed `r length(unique(bk$sample.no))` Bacteria Killing Assays (BKA) using fresh plasma from `r length(unique(bk$Individual_Band))` female breeding TRES. 
We diluted 5 $\mu$L of plasma in 95 $\mu$L of CO2-independent medium. We then diluted it further to a 1.25% plasma-medium dilution. 
Each sample was then challenged with PBS containing `r min(min(bk$positive.1, na.rm = TRUE),min(bk$positive.2, na.rm = TRUE))` to `r max(max(bk$positive.1, na.rm = TRUE),max(bk$positive.2, na.rm = TRUE))` *Escherichia coli* ATCC 8739 (Microbiologics, Epower *E. coli* ATCC 8739) and left to incubate for 45 minutes in a water bath at 40° C.
We then plated 50 $\mu$L of the sample on a petri dish with Tryptic Soy Agar and incubated these for 12h at 37° C. 
We counted all the colonies that grew overnight. 
Each sample was challenged twice and each challenge was plated in two petri dishes. 
We calculated bactericidal ability as the proportion of bacteria killed based on a challenge with no plasma. 
The data in this study is the average of those data counts.

## Assay repeatability

We our data to assess assay repeatability across plates and sample dilutions.

```{r repeatability, fig.width = 8, fig.height = 4}
# Repeatability by plate

pseurep <- bk %>%
  ggplot() + geom_point(aes(x=rep1, y=rep2)) +
  geom_abline(slope =1) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none") +
  labs(title="Pseureplicates - Plate effect",
       x = "No. of CFUs in Plate 1",
       y = "No. of CFUs in Plate 2") +
  scale_y_continuous(expand = c(0,0), limits = c(0,700)) +
  scale_x_continuous(expand = c(0,0), limits = c(0,700)) +
  annotate("text", x = 500, y = 600, 
           label = paste("r","=",
                         round(cor(cbind(bk$rep1, bk$rep2),
                                   method = "pearson")[1,2], 3) ) )

# Repeatability by reaction

reps <- inner_join(filter(bk, rxn.rep == 1) %>% select(Individual_Band, bkc),
           filter(bk, rxn.rep == 2) %>% select(Individual_Band, bkc),
           by = "Individual_Band")

rxnrep <- reps %>%
  ggplot() + geom_point(aes(x=bkc.x, y=bkc.y)) +
  geom_abline(slope =1) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none") +
  labs(title="Reaction repeatibility",
       x = "Prop. Bacteria Killed in Challenge 1",
       y = "Prop. Bacteria Killed in Challenge 2") +
  annotate("text", x = 0.5, y = 0.95, 
           label = paste("r =",round(cor(reps[,-1])[1,2], 3) ) ) 

grid.arrange(pseurep, rxnrep, ncol = 2)
```

## Sample type comparison

Since a subset of plasma samples were taken from blood samples taken 30 minutes after disturbance, we evaluated if there was a difference in BKA between these sample types.

```{r water to land comparison}
# Boxplot
bk %>%
  group_by(Individual_Band, bleed.type) %>%
  ggplot() + geom_boxplot(aes(x = bleed.type, y = bkc, fill=bleed.type), 
                          size = 1, width = 0.25) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none") +
  labs(x = "Sample timing", y = "Prop. Bacteria Killed") +
  scale_x_discrete(labels = c("< 3 minutes", "30 min"))

# T-test
t.test(bkc~ bleed.type, 
       data = bk %>% group_by(Individual_Band, bleed.type) %>% 
         summarise(bkc=mean(bkc))
       ) %>% tidy() %>% kable()
```
# Results

## BKA in our population

```{r bka summary}
bk %>%
  group_by(Individual_Band) %>% 
  summarise(bkc = mean(bkc)) %>%
  ggplot() + geom_histogram(aes(bkc), bins = 15, colour = "white") +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(limits = c(0, 10), breaks=seq(0,10,by=2),
                     expand = c(0,0)) +
  geom_vline(xintercept=0, linetype = "dashed", colour = "grey") +
  labs(x = "Prop. Bacteria Killed", y = "No. of individuals")

summ_bk <- group_by(bk, Individual_Band) %>% summarise(bkc = mean(bkc))
summ <- matrix(0, nrow = 1, ncol = 6)
colnames(summ) <- c("Mean", "S.D.", "2.5%", "97.5%", "Min.", "Max.")
summ[1,1] <- mean(summ_bk$bkc, na.rm = TRUE)
summ[1,2] <- sd(summ_bk$bkc, na.rm = TRUE)
summ[1,3] <- summ[1,1] - 1.96*summ[1,2]/sqrt(nrow(summ_bk))
summ[1,4] <- summ[1,1] + 1.96*summ[1,2]/sqrt(nrow(summ_bk))
summ[1,5] <- min(summ_bk)
summ[1,6] <- max(summ_bk)
kable(summ)
```

## BKA and Reproductive Effort

### General trend evaluation

#### BKA and age

```{r BKA age}
summary(glm(bk ~ Age,  family=Gamma(link="inverse"),
            data = filter(df, Capture_Number == 1),))
```

#### BKA and Lay Date

```{r bka laydate}
summary(glm(bk ~ ylaydate, family = Gamma(link="inverse"),
                  data = filter(df, Capture_Number == 1)))
```

### Clutch Size

#### Clutch size data distribution

```{r clutch distribution}
ggplot(filter(df,Capture_Number == 1) ) +
                 geom_histogram(aes(x=Clutch_Size), 
                                colour = "white", bins = 15) +
                 theme(panel.background = element_rect(fill=NA),
                       axis.line = element_line(size=1)) +
                 scale_y_continuous(expand = c(0,0)) +
                 labs(x= "Clutch size",
                      y = "No. of individuals") +
                 scale_x_continuous(limits = c(0,8),
                                    breaks = 1:8)
```

#### Clutch size models

##### Null model

```{r clutch null}
clutch_m1 <- glm(Clutch_Size ~ Age + ylaydate,
                family = quasipoisson(link = "log"), 
                data = filter(df, Capture_Number == 1))
summary(clutch_m1)
```

##### BKA model

```{r clutch BKA}
clutch_m2 <- glm(Clutch_Size ~ Bacteria_Killing_Assay + Age + ylaydate,
                family = quasipoisson(link = "log"), 
                data = filter(df, Capture_Number == 1))
summary(clutch_m2)
```

##### Interaction model

```{r clutch interaction}
clutch_m3 <- glm(Clutch_Size ~ Bacteria_Killing_Assay*ylaydate + Age,
                 family = quasipoisson(link = "log"), 
                 data = filter(df, Capture_Number == 1))
summary(clutch_m3)
```

##### Model comparison

```{r aic clutch}
# redefine models to poisson models
clutch_m1p <- glm(Clutch_Size ~ Age + ylaydate,
                family = poisson(link = "log"), 
                data = filter(df, Capture_Number == 1))
clutch_m2p <- glm(Clutch_Size ~ Bacteria_Killing_Assay + Age + ylaydate,
                family = poisson(link = "log"), 
                data = filter(df, Capture_Number == 1))
clutch_m3p <- glm(Clutch_Size ~ Bacteria_Killing_Assay*ylaydate + Age,
                 family = poisson(link = "log"), 
                 data = filter(df, Capture_Number == 1))

ICtab(clutch_m1p, clutch_m2p, clutch_m3p, dispersion = dfun(clutch_m1p), type = "qAIC")
```

### Nestling Feeding Rate

#### Feeding Rate variation

```{r visits variation}

# Note that Offset refers to Nestling Age.

ggplot(rfid_d) + 
  geom_point(aes(x= Offset, y = FemFeed, color = f_rfid), show.legend = F) + 
  geom_line(aes(x= Offset, y = FemFeed, color = f_rfid), show.legend = F) +
  scale_x_continuous(expand = c(0,0), limits = c(0,20)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1),
        legend.position = "none") +
  labs(title="Daily provisioning visits per female",
       x = "Nestling Age",
       y = "Number of visits")
```
#### Feeding Rate models

```{r data scaling}
scaled_pc <- pc %>%
  mutate(scBrood_Size_Hatching = scale(Brood_Size_Hatching),
         scOffset = scale(Offset),
         scHour = scale(Hour),
         scjlaydate = scale(jlaydate),
         sctemp = scale(temp),
         scMalFeed = scale(MalFeed),
         scbkc = scale(bkc),
         Individual_Band = as.factor(Individual_Band))
```

##### For all nests

###### Null model

These models are modified from [@vitousekLingeringImpactStress2018]

```{r feeding null}
provisioning_m1 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band),
                         family=poisson(link="log")); 
summary(provisioning_m1)

```

###### BKA model

```{r feeding BKA}
# bkc model
provisioning_m2 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band) + scbkc,
                         family=poisson(link="log"))
summary(provisioning_m2)
```

###### Interaction model

```{r feeding interaction}
provisioning_m3 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                            I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                            scHour + I(scHour^2) + sctemp + 
                            Treatment + scBrood_Size_Hatching*scOffset + 
                            (1|Individual_Band) + scbkc*scjlaydate,
                          family=poisson(link="log"))
summary(provisioning_m3)
```

###### Model comparison

```{r aic provisioning}
ICtab(provisioning_m1, provisioning_m2, provisioning_m3, type = "AIC")
```

#### Including Mate provisioning

##### Null model

```{r mate provisioning null}
provisioning_m4 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scMalFeed + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band),
                         family=poisson(link="log"),
                         control = glmerControl(optimizer = "Nelder_Mead")); 
summary(provisioning_m4)
```

##### BKA model

```{r mate provisioning bka}
provisioning_m5 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + scjlaydate + sctemp + 
                           Treatment + scMalFeed + scBrood_Size_Hatching*scOffset + 
                           (1|Individual_Band) + scbkc,
                         family=poisson(link="log"))
summary(provisioning_m5)
```

##### Interaction model

```{r mate provisioning interaction}
provisioning_m6 <- glmer(data = scaled_pc, FemFeed ~ scBrood_Size_Hatching + 
                           I(scBrood_Size_Hatching^2) + scOffset + I(scOffset^2) + 
                           scHour + I(scHour^2) + sctemp + 
                           Treatment + scBrood_Size_Hatching*scOffset + scMalFeed +
                           (1|Individual_Band) + scbkc*scjlaydate,
                         family=poisson(link="log"))
summary(provisioning_m6)
```

###### Model Comparison

```{r mate provisioning AIC}
ICtab(provisioning_m4, provisioning_m5, provisioning_m6, type = "AIC")
```

### Seasonal Mass Loss

#### Mass loss distribution

```{r mass loss histogram}
df_massloss <- left_join(filter(df, Experiment == "Color_Stressor", Capture_Number == 1),
          dplyr::select(filter(df, Experiment == "Color_Stressor", Capture_Number == 3), Individual_Band, Mass), 
          by = "Individual_Band") %>%
  mutate(prop.ml = (Mass.y - Mass.x)/Mass.x)

ggplot(df_massloss) +
  geom_histogram(aes(x=prop.ml),
                 colour = "white", bins = 15) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x= "Mass lost (g)",
       y = "No. of individuals")
```

#### Mass loss models

##### Null model

```{r mass loss null}
prop_mass_loss_m1 <- glm(prop.ml ~ ylaydate + Brood_Size_Hatching + Treatment2, 
                         family = gaussian(link = "identity"),
                         data = df_massloss)
summary(prop_mass_loss_m1)
```

##### BKA model

```{r mass loss bka}
prop_mass_loss_m2 <- glm(prop.ml ~ Bacteria_Killing_Assay + ylaydate + Brood_Size_Hatching + Treatment2, 
                         family = gaussian(link = "identity"),
                         data = df_massloss)
summary(prop_mass_loss_m2)
```

##### Interaction model

```{r mass loss interaction}
prop_mass_loss_m3 <- glm(prop.ml ~ Bacteria_Killing_Assay*ylaydate + Brood_Size_Hatching + Treatment2, 
                         family = gaussian(link = "identity"),
                         data = df_massloss)
summary(prop_mass_loss_m3)
```

##### Model comparison

```{r aic mass loss}
ICtab(prop_mass_loss_m1, prop_mass_loss_m2, prop_mass_loss_m3, type = "AIC")
```

## BKA and reproductive success

### Fledging success

#### Fleding success models

##### Null model
```{r success null}
success_m1 <- glm(Success ~ Brood_Size_Hatching + ylaydate + Treatment2, 
                  family = binomial(link = "logit"), 
                  data = filter(df, Capture_Number == 1) %>%
                    mutate(Success = ifelse(Nest_Fate == "Fledged", 1, 0)))
summary(success_m1)
```

##### BKA model
```{r success BKA}
success_m2 <- glm(Success ~ Bacteria_Killing_Assay + Brood_Size_Hatching + 
                    ylaydate + Treatment2, family = binomial(link = "logit"), 
                  data = filter(df, Capture_Number == 1) %>%
                    mutate(Success = ifelse(Nest_Fate == "Fledged", 1, 0)))
summary(success_m2)
```

##### Interaction model
```{r success interaction}
success_m3 <- glm(Success ~ Bacteria_Killing_Assay*ylaydate + 
                    Brood_Size_Hatching + Treatment2, 
                  family = binomial(link = "logit"), 
                  data = filter(df, Capture_Number == 1) %>%
                    mutate(Success = ifelse(Nest_Fate == "Fledged", 1, 0)))
summary(success_m3)
```

##### Model comparison
```{r success AIC}
ICtab(success_m1, success_m2, success_m3, type = "AIC")
```

### Number of nestlings fledged

#### Number fledged distribution

```{r fledged histogram}
ggplot(filter(df, Capture_Number == 1, .fledged > 0 )) +
  geom_histogram(aes(x=.fledged),
                 colour = "white", bins = 15) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x= "Number of nestlings fledged",
       y = "No. of individuals")
```

#### Number fledged models

##### Null model

```{r nfledged null}
fledged_m1 <- glm(.fledged ~ Brood_Size_Hatching + ylaydate +  Treatment, 
                  family = quasipoisson(link = "log"), 
                  data = filter(df, Capture_Number == 1, .fledged > 0 ) )
summary(fledged_m1)
```

##### BKA model

```{r nfledged BKA}
fledged_m2 <- glm(.fledged ~ Bacteria_Killing_Assay + Brood_Size_Hatching + 
                    ylaydate + Treatment, family = quasipoisson(link = "log"), 
                  data = filter(df, Capture_Number == 1, .fledged > 0 ) )
summary(fledged_m2)
```

##### Interaction model

```{r nfledged interaction}
fledged_m3 <- glm(.fledged ~ Bacteria_Killing_Assay*ylaydate + 
                    Brood_Size_Hatching + Treatment, 
                  family = quasipoisson(link = "log"), 
                  data = filter(df, Capture_Number == 1, .fledged > 0 ) )
summary(fledged_m3)
```

##### Model comparison

```{r aic nfledged}
# redefine models to poisson
fledged_m1p <- glm(.fledged ~ Brood_Size_Hatching + ylaydate +  Treatment, 
                  family = poisson(link = "log"), 
                  data = filter(df, Capture_Number == 1, .fledged > 0 ) )
fledged_m2p <- glm(.fledged ~ Bacteria_Killing_Assay + Brood_Size_Hatching + 
                    ylaydate + Treatment, family = poisson(link = "log"), 
                  data = filter(df, Capture_Number == 1, .fledged > 0 ) )
fledged_m3p <- glm(.fledged ~ Bacteria_Killing_Assay*ylaydate + 
                    Brood_Size_Hatching + Treatment, 
                  family = poisson(link = "log"), 
                  data = filter(df, Capture_Number == 1, .fledged > 0 ) )

ICtab(fledged_m1p, fledged_m2p, fledged_m3p, dispersion = dfun(fledged_m1p), type = "qAIC")
```

### Nestling mass

#### Nestling mass distribution

```{r nestling mass histogram}
ggplot(filter(nd, Capture_Number == 1, Nestling_Fate == "Fledged")) +
  geom_histogram(aes(x = Nestling_Mass),
                 colour = "white", bins = 15) +
  theme(panel.background = element_rect(fill=NA),
        axis.line = element_line(size=1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x= "Nestling mass at day 12 (g)",
       y = "No. of individuals")
```

#### Nestling mass models

##### Null model

```{r nestling mass null}
nd <- nd %>%
  mutate(Unit_Box = paste(Site, Nest, sep = "_"))


nestling_m1 <- lmer(Nestling_Mass ~ Brood_Size_Hatching + 
                     ylaydate + Treatment2 + temp + (1|Unit_Box), 
                   data = filter(nd, Capture_Number == 1, 
                                 Nestling_Fate == "Fledged",
                                 Bacteria_Killing_Assay != "NA") )
summary(nestling_m1)
```

##### BKA model

```{r nestling mass bka}
nestling_m2 <- lmer(Nestling_Mass ~ Bacteria_Killing_Assay + 
                     Brood_Size_Hatching + temp +
                      ylaydate + Treatment2 + (1|Unit_Box), 
                   data = filter(nd, Capture_Number == 1, 
                                 Nestling_Fate == "Fledged",
                                 Bacteria_Killing_Assay != "NA") )
summary(nestling_m2)
```

##### Interaction model

```{r nestling mass interaction}
nestling_m3 <- lmer(Nestling_Mass ~ Bacteria_Killing_Assay*ylaydate + 
                     Brood_Size_Hatching + temp +
                      Treatment2 + (1|Unit_Box), 
                   data = filter(nd, Capture_Number == 1, 
                                 Nestling_Fate == "Fledged",
                                 Bacteria_Killing_Assay != "NA") )
summary(nestling_m3)
```

##### Model comparison

```{r aic neslting mass}
ICtab(nestling_m1, nestling_m2, nestling_m3, type = "AIC")
```


# References

